Interim report

### 2018-05-31 DraftReport.tex
stationary tests added

\subsection{Rain quantiles}
Since CMIP5 uses earlier years then what we have data from as reference period, it is not possible to compare the numbers with each other, but more the behaviour of the curve. 84-93 is picked as a reference period instead of 83-92 to avoid the clearly lower values in -83. In figure \ref{R95} both curves seems to exhibit a very similar behaviour, which is a steady increase in the total rainfall on days with heavy rainfall. For days with extreme rainfall (figure \ref{R99}), the simulated mean is very close to the data mean, but the spread among the models is still very large. The CMIP5 mean is again showing a steady inscrease which is not clearly visible in the data plot. So similar differences can been seen both when looking at the highest percentiles and very heavy rainfall in mm. 
\begin{figure}[H]
	\centering
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{R95all.pdf}
		\caption{All stations}
	\end{subfigure}%
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{R95GCMall.pdf}
		\caption{CMIP5 all}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{R95GCMmean.pdf}
		\caption{CMIP5 mean}
	\end{subfigure}
	\caption{Total rain amount in days above 95\% threshold for reference period}
	\label{R95}
\end{figure}

@@

>>>>

\subsection{Rain quantiles}
Since CMIP5 uses earlier years then what we have data from as reference period, it is not possible to compare the numbers with each other, but more the behaviour of the curve. 84-93 is picked as a reference period instead of 83-92 to avoid the clearly lower values in -83. To confirm the robustness of these quantile threshold, I ran a bootstrap with 5000 samples of the same size as the decadal samples. After this, I constructed 95\% interval on the threshold values, both with and without forcing the maximum value to be included in the sample. Doing this showed that the 95 percentile value is very robust to wether the largest value is included or not, since the confidence interval does not change at all. The 99 percentile is slightly higher when forcing the largest value to be included (79.4993, 87.7000 compared to 78.80000, 87.60127 for 03-12 decade). The confidence interval is also not as big as one might expect considering it being a very right tail property. The 95\% interval is small (47.8,  50.7  for 03-12 decade), hence we can conclude that the value chosen is robust, and does not depend on the specific data that we have.

In figure \ref{R95} both curves seems to exhibit a very similar behaviour, which is a steady increase in the total rainfall on days with heavy rainfall. For days with extreme rainfall (figure \ref{R99}), the simulated mean is very close to the data mean, but the spread among the models is still very large. The CMIP5 mean is again showing a steady inscrease which is not clearly visible in the data plot. So similar differences can been seen both when looking at the highest percentiles and very heavy rainfall in mm. 
\begin{figure}[H]
	\centering
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{R95all.pdf}
		\caption{All stations}
	\end{subfigure}%
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{R95GCMall.pdf}
		\caption{CMIP5 all}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{R95GCMmean.pdf}
		\caption{CMIP5 mean}
	\end{subfigure}
	\caption{Total rain amount in days above 95\% threshold for reference period}
	\label{R95}
\end{figure}

@@

	\section{Extreme analysis}
	\subsection{Ground work}
	In order to use the extreme value theory, in our case \textbf{POT} method, we nee our data to be stationary, to ensure identical, and independent. To make data behave independent, we remove the smaller of the two esceedances, if they are closer then a certain number of days from each other. This lag is usually set to 3 or 5 days, depending on region and time of year. To make sure that the exceedances follows a \textbf{Poisson point process},the standard way is to consider data in cluster of a specific size and then only take the maximum if more then one observation within that cluster exceeds the threshold. This makes it important to pick a high enough threshold, since else a lot of our observations will be lost.
	To ensure stationarity is more diffucult and several approaches are used. One is to divide the data into seasons and use different thresholds for each season. The issue with this method is that we assume stationarity within each season and therefore can get unrealistic return values if there in fact is a trend in the data. If one knows that the data depends on other factors, such as the yield depends on temperature and rain, one can use a Box-Cox transformation on the data and then let the drift and trend parameter depend on these other paramters, so called \textbf{covariates}. This however requires that we know of these covariates, can estimate or already have models for how they affect our data and have measurments of these covariates over the same time period as our own data.\\

	{\Large\textbf{HO}}: \textbf{MAM}: Both trend tests gives high p-values, i.e no evidence for a trend in the time series. Both unit root tests indicates that the series is stationary/ trend stationary. No significant autocorrelation or pattern in lag 1 plot.
	\textbf{JJA}: Using p=0.1 we reject the null hypothesis and conclude that we have a decresing trend, but not on a 5\% level (MK does not confirm). Same result as above for unit root and autocorrelation.
	\textbf{SON}: No evidence for a trend, unit root or autocorrelation. 
	\textbf{DJF}: Same as above.
	None of the stationary seasons rejects the null hypothesis about rough truncation, hence we can use the normal POT method (?). The shape parameter for DJF is estimated to be -0.1. MAM is estimated to have a shape parameter 0.2 and SON 0.\\

	{\Large\textbf{KRA}}: \textbf{MAM} : Same as above. \textbf{JJA}: Decreasing trend on a 10\% level (MK confirms). No evidence of a unit root or autocorrelation. \textbf{SON}: Strong evidence for an increasing trend. Significant autocorrelation for up tp 3 day lag. \textbf{DJF}: Evidence on a 10\% level for an increasing trend.No autocorrelation. \\

	{\Large\textbf{ADA}}: \textbf{MAM}: Same as above. \textbf{JJA}: Evidence on a 5\% level for an incresing trend. No evidence for a unit root, but significant autocorrelation up to 11 days lag and some linear pattern in plot of 1 day lag. \textbf{SON}: Strong evidence for an increasing trend, but not for a unit root. No significant autocorrelation. \textbf{DJF}: Evidence on 5\% level for an increase, but no evidence for unit root or autocorrelation.
